---
title: "Untitled"
output: pdf_document
---


```{r echo = T, results = 'hide', error=FALSE, warning=FALSE, message=FALSE}
load("C:/Users/emrek/OneDrive/Desktop/bekaya-dataessay/raw-data/pko.RData")
```

```{r}
head(pko)
dim(pko)
summary(pko)
table(is.na(pko)) #60665 
length(table(pko$conflict_ID)) #84 different conflict_ID
pko[pko$brf_grc == max(na.omit(pko$brf_grc)),] #the highest fatality belongs to the conflict between Congo and Cobras.
table(pko$ceasefire) #0:12452  1:9066
pko[pko$pop_ln == min(na.omit(pko$pop_ln)),] #Comoros has the smallest population size
table(pko$biased_intervention) #1: 744 
table(pko$active_year) #1: 4034
table(pko$regional_pko)


```

```{r}
#Data preperation

#Creating a new variable for the sum of Number of UN military troops and Number of UN police unites deployed, in thousands, lagged one month.

pko$military_police_lag1000 <- pko$troop_lag1000 + pko$police_lag1000

#Creating a dataset by using the variables I want to focus
#Brf_grc for battle-related fatalities, ceasefire, biased_interventation, active_year and military_police_lag1000 for UN troops size variables 
pko2 <- pko[,c(5,9,16,17,22)]
summary(pko2)
dim(pko2) #21518     
pko2omit <- na.omit(pko2)
summary(pko2omit)
dim(pko2omit) #From 21518 to 19792 observations.


mean(pko2omit[pko2omit$ceasefire==0,]$brf_grc) #10
mean(pko2omit[pko2omit$ceasefire==1,]$brf_grc) #4

mean(pko2omit[pko2omit$ceasefire==0,]$military_police_lag1000) #0.84
mean(pko2omit[pko2omit$ceasefire==1,]$military_police_lag1000) #2
```


```{r}
library(ggplot2)
library(viridisLite)
library(viridis)

# to see the relationship between UN troops size and battle-related fatalities by ceasefire
col_vec <- viridis(2, alpha = 0.5)

# Why do we nest two ifelse statements?
attend_lab <- ifelse(pko2omit$ceasefire == 1, 
                     "Ceasefire exists",
                     ifelse(pko2omit$ceasefire == 0, 
                            "Ceasefire doesn't exist", NA)) 

# instead of nesting two ifelse statements, we can also use case_when form dplyr 
# attend_lab <- case_when(trump_rally$attendance == 1 ~ "did attend",
#                         trump_rally$attendance == 0 ~ "did not attend",
#                         TRUE ~ NA) # assign NA if trump_rally$attendance is not 1 or 0

ggplot(data = pko2omit,  # data used for plotting
       mapping = aes(x = pko2omit$military_police_lag1000, 
                     y = pko2omit$brf_grc,
                     color = attend_lab)
       ) +
  geom_jitter(
    size = 2,
    width = 0.25
       ) + 
  theme_minimal() + # change the appearance
  labs(
    x = "The UN troops size",
    y = "Battle-related fatalities",
    title = ""
      ) + 
  scale_color_manual(
    name = "", 
    values = c("Ceasefire exists" = col_vec[1], "Ceasefire doesn't exist" = col_vec[2])
    )
```

```{r}
#to check the correlations between the variables in order to have insights 

cor(pko2omit$brf_grc,pko2omit$military_police_lag1000) #With the correlation=-0.01871248, it can be said that there is a weak negative relationship between Battle-related fatalities and the UN troops size. 

data1 = subset(pko2omit, ceasefire == 1) #ceasefire exists
data2 = subset(pko2omit, ceasefire == 0) #ceasefire doesn't exist

cor(data1$military_police_lag1000, data1$brf_grc)  #-0.019

cor(data2$military_police_lag1000, data2$brf_grc)  #-0.0107

#Interestingly, when the ceasefire existss, there is no big change but the negative relationship gets a very little stronger. On the other hand, when there is no ceasefire between groups, the negative relationship between the groups gets weaker from -0.0187 to -0.0107.
```

```{r}
library(stargazer)
#Poisson
poissonmodel1 <- glm(
  brf_grc ~ military_police_lag1000, 
  data = pko2omit,
  family = "poisson"
)

poissonmodel2 <- glm(
  brf_grc ~ military_police_lag1000 + ceasefire + military_police_lag1000 * ceasefire , 
  data = pko2omit,
  family = "poisson"
)

#negativebinomial
library(MASS)
nbmodel1 <- glm.nb(
  brf_grc ~ military_police_lag1000, 
  data = pko2omit,
  control = glm.control(maxit = 100)
)

nbmodel2 <- glm.nb(
  brf_grc ~ military_police_lag1000 + ceasefire,
  data = pko2omit,
  control = glm.control(maxit = 100)
)

nbmodel3 <- glm.nb(
  brf_grc ~  military_police_lag1000 +ceasefire + biased_intervention + active_year, 
  data = pko2omit,
  control = glm.control(maxit = 100)
)

nbmodel4 <- glm.nb(
  brf_grc ~ military_police_lag1000 +ceasefire + biased_intervention + active_year  + ceasefire * military_police_lag1000 , 
  data = pko2omit,
  control = glm.control(maxit = 200)
)

stargazer(nbmodel1,nbmodel2,nbmodel3,nbmodel4,
          out = "table_lab.tex",
          title="Regression Results", 
          notes = "Evaluating UN Peacekeeping Operations",
          intercept.bottom = TRUE,
          align=TRUE, 
          covariate.labels=c("Number of UN troops","Ceasefire", "Biased Intervention","Active Year",
          "Number of UN Troops : Ceasefire"),
          no.space=TRUE,
          dep.var.labels = c("Battle-related fatalities"),
          table.placement = "H",
          type="text")
```
Akaike Inf. Crit. : The Akaike information criterion (AIC) is a mathematical method for evaluating how well a model fits the data it was generated from. The smaller the AIC value, the better the model fit.

```{r}
library(car)
vif(nbmodel4) #multicollinearity check - doesn't exist
```
#Quantities of Interest

Steps for Simulating Parameters (Estimation Uncertainty):

Get the coefficients from the regression (gamma_hat)
Get the variance-covariance matrix (V_hat)
Set up a multivariate normal distribution N(gamma_hat,V_hat)
Draw from the distribution nsim times

```{r}
nsim <- 1000

# 1. get the coefficients
gamma_hat <- coef(nbmodel4)

# 2. Get the variance-covariance matrix
V_hat <- vcov(nbmodel4)

# 3. Set up a multivariate normal distribution N(gamma_hat, V_hat)
# 4. Draw from the distribution nsim times
S <- mvrnorm(nsim, gamma_hat, V_hat)

```

Next step: Set up interesting scenarios.

The order of coefficients is:

```{r}
names(gamma_hat)
```

```{r}
# create sequence for the continuous variable
troops_seq <- seq(
  min(nbmodel4$model$military_police_lag1000),
  quantile(nbmodel4$model$military_police_lag1000, 0.95), 
  length.out = 100
)

# set scenario
scenario0 <- cbind(
  1, # Intercept
  troops_seq, # mean of military_police_lag1000
  0, #ceasefire
  mean(pko2omit$biased_intervention), #biased_intervention
  1, #active year
  troops_seq * 0 # military_police_lag1000:ceasefire
)

colnames(scenario0) <- names(gamma_hat)
head(scenario0)

```


```{r}
scenario1 <- scenario0

# switch only the changing values
# set scenario for ceasefire=1
scenario1[, which(colnames(scenario1) == "ceasefire")] <- 1
scenario1[, which(colnames(scenario1) == "military_police_lag1000:ceasefire")] <- troops_seq
head(scenario1)
```


```{r}
Xbeta0 <- S %*% t(scenario0)
Xbeta1 <- S %*% t(scenario1)

lambda0 <- exp(Xbeta0)
lambda1 <- exp(Xbeta1)
```


```{r}
theta <- nbmodel3$theta

exp_0 <-
  apply(lambda0, c(1, 2), function(x) {
    mean(rnbinom(100, size = theta, mu = x))
  })

exp_1 <-
  apply(lambda1, c(1, 2), function(x) {
    mean(rnbinom(100, size = theta, mu = x))
  })
```


```{r}
quants_0 <- t(apply(exp_0, 2, quantile, c(0.025, 0.5, 0.975)))
quants_1 <- t(apply(exp_1, 2, quantile, c(0.025, 0.5, 0.975)))
```

```{r}
par(las = 1)
# segment plot-

plot(
  troops_seq,
  quants_0[, "50%"],
  type = "n",
  ylim = c(0, 150),
  ylab ="Battle-related Fatalities",
  xlab ="The UN Troops size",
  bty = "n",
  main = "Expected Battle-related Fatalities 
  over the range of The UN Troops size by Ceasefire",
) 

segments(
  x0 = troops_seq, x1 = troops_seq,
  y1 = quants_0[, "97.5%"], y0 = quants_0[, "2.5%"],
  col = viridis(2, 0.5)[1],
  lwd = 2
)
points(troops_seq, quants_0[, 2], col = viridis(2, 0.5)[1], pch = 20)

segments(
  x0 = troops_seq, x1 = troops_seq,
  y1 = quants_1[, "97.5%"], y0 = quants_1[, "2.5%"],
  col = viridis(2, 0.5)[2],
  lwd = 2
)
points(troops_seq, quants_1[, 2], col = viridis(2, 0.5)[2], pch = 20)

# Add a "histogram" of actual X1-values.
axis(
  1,
  at = pko2omit$military_police_lag1000,
  col.ticks = "gray30",
  labels = FALSE,
  tck = 0.02
)

legend(
  "topright",
  legend = c(
    "Median & 95% CI:",
    "Ceasefire doesn't exit",
    "Ceasefire exists",
    ""
  ),
  col = c(
    "white",
    viridis(2, 0.5)
  ),
  lty = "solid",
  lwd = 2,
  pch = 20,
  pt.cex = 2,
  bty = "n"
)

```

```{r}
# polygon plot
plot(
  troops_seq,
  quants_0[, "50%"],
  type = "n",
  ylim = c(0, 150),
  ylab ="Battle-related Fatalities",
  xlab ="The UN Troops size",
  bty = "n",
  main = "Expected Battle-related Fatalities 
  over the range of The UN Troops size by Ceasefire",
) 

polygon(
  c(rev(troops_seq), troops_seq),
  c(rev(quants_0[, "97.5%"]), quants_0[, "2.5%"]),
  col = viridis(2, 0.2)[1],
  border = NA
)

polygon(
  c(rev(troops_seq), troops_seq),
  c(rev(quants_1[, "97.5%"]), quants_1[, "2.5%"]),
  col = viridis(2, 0.2)[2],
  border = NA
)

lines(troops_seq, quants_0[, 1], lwd = 2, lty = "dashed", col = viridis(2, 0.5)[3])
lines(troops_seq, quants_0[, 2], lwd = 0.5, lty = "dashed", col = viridis(2, 0.5)[3])

lines(troops_seq, quants_1[, 1], lwd = 2, lty = "dotted", col = viridis(2, 0.5)[2])
lines(troops_seq, quants_1[, 2], lwd = 0.5, lty = "dashed", col = viridis(2, 0.5)[2])

# Add a "histogram" of actual X1-values.
axis(
  1,
  at = pko2omit$military_police_lag1000,
  col.ticks = "gray30",
  labels = FALSE,
  tck = 0.02
)

legend(
  "topright",
  legend = c(
    "Median & 95% CI:",
    "Ceasefire doesn't exit",
    "Ceasefire exists"
  ),
  col = c(
    "white",
    viridis(2, 0.5)
  ),
  lty = c(NA, "dashed", "dotted"),
  lwd = c(NA, 2, 2, 2),
  # pch = c(NA, 15, NA, 15, NA, 15),
  pt.cex = 2,
  bty = "n"
)
```

```{r}
FD <- exp_0 - exp_1
quants_FD <- t(apply(FD, 2, quantile, c(0.025, 0.5, 0.975)))

plot(
  troops_seq,
  quants_FD[, "50%"],
  type = "n",
  ylim = c(min(quants_FD[, "2.5%"]), max(quants_FD[, "97.5%"])),
  ylab = "Difference in Expected Battle-related Fatalities",
  xlab = "The UN Troops Size",
  bty = "n",
  las = 1,
  main = "Difference in expected number of Battle-related Fatalities 
  over the range of The UN Troops size by Ceasefire",
)
segments(
  x0 = troops_seq, x1 = troops_seq,
  y1 = quants_FD[, "97.5%"], y0 = quants_FD[, "2.5%"],
  col = viridis(1, 0.5)
)
points(troops_seq, quants_FD[, 2], col = viridis(1, 0.5), pch = 20)

abline(h = 0, lty = "dashed")

```

```{r}
FD <- cbind(
  exp_0[, 97.5] - exp_0[, 2.5],
  exp_1[, 97.5] - exp_1[, 2.5]
)

quants_FD <- t(apply(FD, 2, quantile, c(0.025, 0.5, 0.975)))

plot(
  y = 1:2,
  x = quants_FD[, "50%"],
  type = "n",
  xlab = "Difference in Expected Battle-related Fatalities",
  ylab = "",
  xlim = range(pretty(c(
    min(quants_FD[, "2.5%"]), max(quants_FD[, "97.5%"])
  ))),
  ylim = c(1, 2.6),
  bty = "n",
  las = 1,
  axes = F,
  main = "Difference in expected number of Battle-related Fatalities 
  of The UN Troops by Ceasefire"
)

segments(
  x0 = quants_FD[, "2.5%"],
  x1 = quants_FD[, "97.5%"],
  y0 = c(1.2, 1.8),
  col = viridis(5, 0.5)
)
points(
  x = quants_FD[, "50%"],
  y = c(1.2, 1.8),
  pch = 19
)
axis(
  2,
  at =  c(1.2, 1.8),
  las = 1,
  labels = c("Ceasefire=0", "Ceasefire=1"),
  tick = F,
  line = F,
  hadj = 0.23
)
axis(1)
abline(v = 0, lty = "dashed")
```

```{r}
#Robustness Check

#Outlier elimination
pkomodel <- nbmodel4 <- glm.nb(
  brf_grc ~  ceasefire + military_police_lag1000 + active_year +  biased_intervention  + military_police_lag1000 * ceasefire, 
  data = pko2omit,
  control = glm.control(maxit = 100)
)

pkomodel2 <- glm.nb(brf_grc ~  ceasefire + military_police_lag1000 + active_year + biased_intervention  + military_police_lag1000 * ceasefire, 
        data = pkomodel$model[!rstandard(pkomodel) > 2, ], control = glm.control(maxit = 200), )

stargazer(pkomodel,pkomodel2,
          out = "table_lab.tex",
          title="Regression Results", 
          notes = "Evaluating UN Peacekeeping Operations",
          intercept.bottom = TRUE,
          align=TRUE, 
          covariate.labels=c("Ceasefire", "Number of UN troops", "Active Year", "Biased Intervention",
          "Number of UN Troops : Ceasefire"),
          no.space=TRUE,
          dep.var.labels = c("Battle-related fatalities"),
          table.placement = "H",
          type="text")

summary(pkomodel)
```






